### 2016-09-16 Slides/morality-slides.Rmd
First commit.

---
title: "The Morality of Arbitrage"
author: "Tyler J. Brough"
date: "September 16, 2016"
output: pdf_document
---

#

> But nobody can be a great economist who is only an economist - and I am even tempted to add that the economist who
> is only an economist is likely to become a nuisance if not a positive danger.

-- Friedrich von Hayek