### 2016-09-16 Slides/morality-slides.Rmd
Progress on slides.

---
title: "The Morality of Arbitrage"
author: "Tyler J. Brough"
date: "September 16, 2016"
output: ioslides_presentation
---

#

> *But nobody can be a great economist who is only an economist - and I am even tempted to add that the economist who is only an economist is likely to become a nuisance if not a positive danger*.

-- Friedrich von Hayek

## Introduction

>- I have been very influenced by Coases's 1975 AER article *The Market for Goods and the Market for Ideas*.
>- I have also been reading articles lately by Robert Nau and his coauthors. In particular the paper by Nau & McCardle (1991) *Aribtrage, Rationality, and Equilibrium*. 
>- This project is an attempt to do something like what Coase had in mind, using the the approach of Nau, but applied to the morality of markets and capitalism.

## Nau's Arbitrage Choice Theory