### 2017-04-22 Draft/draft2.Rmd
Adding draft2.

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ./svm-latex-ms.tex
title: "The Morality of Arbitrage"
author:
- name: Tyler J. Brough
  affiliation: Utah State University
abstract: "In this paper, I argue that one cannot rationally believe that financial arbitrage is an inherently immoral pursuit, while simultaneously believing that the pursuit of academic knowledge is an inherently moral pursuit. I argue that these two seemingly separate activities are really one and the same. Further those that do hold such contradictory beliefs are behaving incoherently and have a Dutch book made against them."
keywords: "arbitrage, Dutch books, morality"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: ./master.bib
biblio-style: apsr
---

# Introduction

> [T]he shipper who earns his living from using otherwise empty or half-filled journeys of tramp-steamers, or the estate
> agent whose whole knowledge is almost exclusively one of temporary opportunities, or the *arbitrageur* who gains from
> local differences of commodity prices, are all performing eminently useful functions based on special knowledge of
> circumstances of the fleeting moment not known to others.
	-- F.A. Hayek

In this essay I follow the lead of @Coase1974, and examine the very common differential treatment of the market for goods and the market for ideas.

The plan of this paper is as follows: section 2 introduces Coases's original argument, and his conclusion that
their different treatment is incongruous; section 3 discusses the view in @Hayek1945 that the essential
nature of the economic problem is one of the utilization of knowledge. For Hayek, the economic process is an
information process and markets can be seen, using the modern parlance, as information systems. There is thus
no justification for the differential treatment of goods and ideas; section 4 introduces the concept of
subjective probability following @deFinetti1937, as well as the Arbitrage Choice Theory outlined in @Nau1999;
section 5 explains, as essential background, the Dutch Book Theorem; section 6 gives a few chosen examples of
scholarly debates framed as Dutch Book arguments; section 6 frames Coases's original argument as a Dutch Book
argument; section 7 concludes.

# The Market for Goods and the Market for Ideas

# The Use of Knowledge in Society

# Arbitrage Choice Theory

# The Dutch Book Theorem Explained

## Examples of Dutch Book Arguments

# Coase's Dutch Book Argument

# Conclusion

<!--
# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent
-->